\documentclass[letterpaper,9pt,twocolumn,twoside,]{pinp}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% pinp change: the geometry package layout settings need to be set here, not in pinp.cls
\geometry{layoutsize={0.95588\paperwidth,0.98864\paperheight},%
  layouthoffset=0.02206\paperwidth, layoutvoffset=0.00568\paperheight}

\definecolor{pinpblue}{HTML}{185FAF}  % imagecolorpicker on blue for new R logo
\definecolor{pnasbluetext}{RGB}{101,0,0} %



\title{A Template for Two or One-Column Vignettes}

\author[a]{Luke Gallagher}
\author[a]{Second Author}
\author[a]{Second Author}
\author[a]{Second Author}

  \affil[a]{The University of Sydney}

\setcounter{secnumdepth}{0}

% Please give the surname of the lead author for the running footer
\leadauthor{Gallagher, Kunal,}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 

\begin{abstract}
Using the dataset of housing prices in New York city that identifies the
price and the distinguishing features of each home, we hoped to find out
whether there is an accurate way to predict the price of a home based on
its specific characteristics. It was found through using a combination
of simple regression, comparison of a backward and forward model and by
comparing the performance of a multivariate regression model to that of
the simple regression model that the variables with the highest
correlation to housing prices in New York and therefore serve as the
best predictors of Price are Living.Area, Land.Value and Bathrooms with
Living.Area having the best rate for prediction of overall Price.
\end{abstract}

\dates{This version was compiled on \today} 


% initially we use doi so keep for backwards compatibility
% new name is doi_footer
\doifooter{\url{https://cran.r-project.org/package=YourPackage}}

\pinpfootercontents{YourPackage Vignette}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{This template package builds upon, and extends, the work of the
excellent \href{https://cran.r-project.org/package=rticles}{rticles}
package, and both packages rely on the
\href{http://www.pnas.org/site/authors/latex.xhtml}{PNAS LaTeX} macros.
Both these sources are gratefully acknowledged as this work would not
have been possible without them. Our extensions are under the same
respective licensing term
(\href{https://www.gnu.org/licenses/gpl-3.0.en.html}{GPL-3} and
\href{https://www.latex-project.org/lppl/}{LPPL (\textgreater= 1.3)}).}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\begin{itemize}
\item
  The USA was hit hard by the COVID pandemic and New York City in
  particular was the hardest hit city worldwide.
\item
  Hence, the New York real estate market is a buyer's market with total
  sales to listing ratio of 0.12 i.e.~the supply of homes is much higher
  than the demand for homes. Hence, it is a buyer's market.
\end{itemize}

\[\text{What set of factors best predicts price in a linear model?}\]

\hypertarget{data-set}{%
\subsection{Data set}\label{data-set}}

Per common academic best practice, you can include your department,
institution, and complete address, with the ZIP/postal code, for each
author. Use lower case letters to match authors with institutions, as
shown in the example. Authors with an ORCID ID may supply this
information at submission.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{discussion-and-conclusion}{%
\subsection{Discussion and conclusion}\label{discussion-and-conclusion}}

\hypertarget{references}{%
\subsection{References}\label{references}}

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use \texttt{\textbackslash{}citet\{\}},
\texttt{\textbackslash{}citep\{\}}, etc as usual.

We default to the \texttt{jss.bst} style. To switch to a different
bibliography style, please use \texttt{biblio-style:\ style} in the YAML
header.

%\showmatmethods
\showacknow


\bibliography{pinp}
\bibliographystyle{jss}



\end{document}
